\documentclass[10pt]{article}
\usepackage[table,xcdraw]{xcolor}
\usepackage[margin=0.5in]{geometry}
\begin{document}

% Please add the following required packages to your document preamble:
% 
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[]
\centering
\caption{My understanding of Downscaling}
\label{my-label}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
                     & Scale  & Degree    & Train(have truth) & Validation(have truth, pretend not known) & Test(no truth) \\ \hline
\rowcolor[HTML]{EFEFEF} 
GCM                  & coarser & 2.5*2.5   & has data          & has data                                  & has data       \\ \hline
\rowcolor[HTML]{EFEFEF} 
Reanalysis           & coarser & 1.5*1.5   & has data          & has data                                  & has data       \\ \hline
\rowcolor[HTML]{656565} 
Observation(station) & finer  & 0.25*0.25 & has data          & has data                                  & no data        \\ \hline
\end{tabular}
\end{table}

\begin{itemize}
\item Step 1(model building): 1 coarse + (only) 1 fine = model
\item Step 2(prediction): model + (future) 1 coarse (as input) = downscaling fine 
\item Question: what is the right(suggested) way to do downscaling? 
\item Question: what is the wrong way to do downscaling?
\item Question: what is the not wrong way(but not very good way) to do downscaling?
\begin{itemize}
\item How to choose the coarse scale variable, GCM or Reanalysis? In Step 1 and Step 2.
\item Train + Validation: what is the coarse scale variable, GCM or Reanalysis?
\item Train + Test: what is the coarse scale variable, GCM or Reanalysis?
\end{itemize}
\end{itemize}

Here are the T.J.'s comments

\begin{itemize}
\item[1] \begin{quotation}
For testing downscaling models I'd recommend using a reanalysis dataset rather than a GCM to allow for better evaluation.  When you evaluate using a GCM assumptions are made that the GCM is analogous to the "true" climate, along with others.
\end{quotation}
\item[2] \begin{quotation}
cannot evaluate downscaled GCM projections with observed data
\end{quotation}
\end{itemize}


\begin{table}[]
\centering
\caption{Which one is right/wrong/not suggested?}
\label{my-label2}
\begin{tabular}{|l|l|l|l|}
\hline
       & Train           & Validation      & Test            \\ \hline
Coarser & GCM?Reanalysis? & GCM?Reanalysis? & GCM?Reanalysis? \\ \hline
Finer   & Observation     & Observation     & [Goal: Observation]     \\ \hline
\end{tabular}
\end{table}


So,
\begin{itemize}
\item[1] Train(Reanalysis)+Validation(Reanalysis)+Test(GCM) is right(suggested)?
\item[2] What about: Train(Reanalysis)+Validation(Reanalysis)+Test(Reanalysis)?
\item[3] Train(GCM)+Validation(GCM) is wrong?
\item[4] Train(GCM)+Validation(Reanalysis) is right but not suggested?
\end{itemize}



\end{document}